A Convolutional Neural Network for Outsole Recognition\tnoteref{t1}

### 2020-04-03 paper/paper-revision.Rnw
Add citations, clean up fix tags

\begin{abstract}
To assess random match probability at the level of class characteristics, one could imagine sampling shoes from a local population and calculating the proportion of shoes "similar" to a given print. One implemention of such a procedure could involve 1) a scanner to gather images of shoes and 2) software to automatically label images within a feature space and assess the images for similarity \fix{within} those features. In this paper, we discuss an algorithm for automatically labeling shoe outsole images using the same feature space examiners use: a set of features based on geometric shapes such as circles, stars, and quadrilaterals. The automatic labeling algorithm performs well on images with unambiguous shapes and moderate color contrast, though some systematic errors arise from the physical similarities of geometric shapes. Additional improvements may be realized by modifying the labeling scheme to reduce feature overlap and by preprocessing the images to improve contrast. This new approach represents a significant improvement to the current manual classification of footwear tread patterns, and could be integrated into current matching databases to reduce the amount of time examiners spend on data entry.
\end{abstract}

@@

In the United States, forensic analysis of shoe print evidence in criminal cases is varied and difficult. The process of collecting print evidence from a crime scene is not standardized, and it is often impossible to conduct an individualizing analysis when a print is partial or of low quality. In fact, some experts estimate that the majority of cases in the United States do not consider individualizing characteristics of shoeprint evidence \citep{smith_2011}. %\footnote{Leslie Hammer, presentation to CSAFE on March 5, 2018}).
Even so, in these cases it still may be \mt{both possible and useful} to assess the random match probability of the footwear evidence \mt{at the level of class characteristics}. In other words, given a questioned shoe print, what is the probability that a randomly selected shoe in the relevant population has a similar tread pattern? Note that this question is distinct from make and model determination since many shoes of different makes and models can still have similar tread patterns (e.g., knockoffs, similar models over different years, etc).

@@

However, studies by \fix{authors (yr)} and \fix{authors (yr)} demonstrate that even \fix{[a couple hundred]} shoe prints constitute an unweildy amount of data requiring \fix{[a large amount of data processing]}. Thus, in parallel with the development of such a scanner, it is necessary to develop a way to process large amounts of data automatically. \mt{The primary focus of this paper is the development of one such method for large-scale automatic data processing, where "data processing" refers to assessing} the similarity of the shoe tread patterns in the collected data.

>>>>

\begin{abstract}
To assess random match probability at the level of class characteristics, one could imagine sampling shoes from a local population and calculating the proportion of shoes "similar" to a given print. Our proposed implemention of such a procedure involves 1) a scanner to gather images of shoes and 2) software to automatically label images within a feature space and assess the images for similarity among those features. In this paper, we discuss an algorithm for automatically labeling shoe outsole images using a feature space similar to what examiners use: a set of features based on geometric shapes such as circles, stars, and quadrilaterals. The automatic labeling algorithm performs well on images with unambiguous shapes and moderate color contrast, though some systematic errors arise from the physical similarities of geometric shapes. Additional improvements may be realized by modifying the labeling scheme to reduce feature overlap and by preprocessing the images to improve contrast. This new approach represents a significant improvement to the current manual classification of footwear tread patterns, and could be integrated into current matching databases to reduce the amount of time examiners spend on data entry.
\end{abstract}

@@

In the United States, forensic analysis of shoe print evidence in criminal cases is varied and difficult. The process of collecting print evidence from a crime scene is not standardized, and it is often impossible to conduct an individualizing analysis (i.e., one based on randomly acquired characteristics) when a print is partial or of low quality. In fact, the majority of footwear examinations in the United States reach conclusions regarding the correspondence of class characteristics rather than of individualizing characteristics \citep{smith_2011}. %\footnote{Leslie Hammer, presentation to CSAFE on March 5, 2018}).
In cases where individualizing characteristics are not used, it still may be both possible and useful to assess the random match probability of the footwear evidence at the level of class characteristic; in other words, given a questioned shoe print, what is the probability that a randomly selected shoe in the relevant population has a similar tread pattern? Note that this question is distinct from make and model determination since many shoes of different makes and models can still have similar tread patterns (e.g., knockoffs, similar models over different years, etc).

@@

However, studies by \citet{parent_sandy_significance_nodate} and \citet{benedict_geographical_2014} demonstrate that even \fix{[a couple hundred]} shoe prints constitute an unweildy amount of data requiring \fix{[a large amount of data processing]}. Thus, in parallel with the development of such a scanner, it is necessary to develop a way to process large amounts of data automatically. \mt{The primary focus of this paper is the development of one such method for large-scale automatic data processing, where "data processing" refers to assessing} the similarity of the shoe tread patterns in the collected data.
